---
title: "cleaning outcome data"
author: "Michelle.VanTieghem"
date: "7/23/2020"
output: html_document
---


```{r, include = F, echo = F}
library(tidyverse)

```

# CBCL + clinical outcomes

## load outcome datasets 
```{r}
# binary variables for each diagnosis possible
dx_df <- read.csv("../../data/HBN_Data/Diagnosis.csv")
head(dx_df)

dx_df <- dx_df %>%
  mutate(complete_dx = ifelse(rowSums(is.na(.)) > 0, 0, 1))

# for each subject, how many diagnoses they have - broken into primary, secondary, etc. 
# and then what category and type of diagnosis 
consensus_df <- read.csv("../../data/HBN_data/ConsensusDX_r8.csv")
head(consensus_df)

# get rid of the zillions of columns of mostly NAs 
# just keep primary and secondary DX 
consensus_df <- consensus_df %>%
  select(URSI, contains("01"), contains("02")) %>%
  # mark incomplete evals as NOT complte 
  mutate(complete_con_dx1 = ifelse(DX_01_Cat == "No Diagnosis Given: Incomplete Eval", 0, 1), 
         complete_con_dx2 = ifelse(DX_02_Cat == "No Diagnosis Given: Incomplete Eval", 0, 1))

```

## load cluster input, CBCL

```{r}
cbcl_df <- read.csv("../../data/HBN_data/CBCL_r8.csv")
head(cbcl_df)

cbcl_df <- cbcl_df %>%
  mutate(complete_cbcl = ifelse(rowSums(is.na(.)) > 0, 0, 1)) 
```

## merge together
```{r}
clinical_df <- cbcl_df %>%
  left_join(., dx_df, by = "URSI", all = T) %>%
  left_join(., consensus_df, by = "URSI", all = T) %>%
  mutate(complete_dx = ifelse(is.na(complete_dx), 0, complete_dx), 
         complete_con_dx1 = ifelse(is.na(complete_con_dx1), 0, complete_con_dx1))

```


## missingness tables
cbcl is missing a lot of data
```{r}
clinical_df %>%
  group_by(complete_cbcl) %>%
  count()
```


diagnosis missingness
```{r}
clinical_df %>%
  group_by(complete_dx) %>% 
  count()
```

consensus DX missingness

```{r}
clinical_df %>%
  group_by(complete_con_dx1) %>% 
  count()
```

## of complete CBCL, how many with DX data?
```{r}
clinical_df %>%
  filter(complete_cbcl == 1) %>%
  group_by(complete_dx, complete_con_dx1) %>%
  count()
```


# IQ and cognitive measures
clustering: WASI + WIAT 
outcome: WISC + WAIS

## load outcome datasets 
```{r}

wisc_df <- read.csv("../../data/HBN_data/WISC_r8.csv")
head(wisc_df)

wais_df <- read.csv("../../data/HBN_data/WAIS_r8.csv")
head(wais_df)
```

## load cluster input data 
```{r}
# only use WIAT Complete & WIAT Valid 
wiat_df <- read.csv("../../data/HBN_Data/WIAT_r8.csv")
head(wiat_df)

wasi_df <- read.csv("../../data/HBN_Data/WASI_r8.csv")
head(wasi_df)
```

## merge together
```{r}
cog_df <- wiat_df %>%
  left_join(wasi_df, by = "URSI", all = T) %>%
  left_join(wais_df, by = "URSI", all = T) %>%
  left_join(wisc_df, by = "URSI", all = T) %>%
  mutate(WASI_Complete = replace_na(WASI_Complete, 0), 
         WIAT_Complete = replace_na(WIAT_Complete , 0),
         WAIS_Complete = replace_na(WAIS_Complete, 0),
         WISC_complete = replace_na(WISC_complete, 0))

```

## count missingness
clustering input: we shouldn't use WASI! very small N 
```{r}
cog_df %>%
  group_by(WASI_Complete, WIAT_Complete) %>%
  count()

```

outcomes: we shouldn't use WAIS! very small N 
```{r}
cog_df %>%
  group_by(WISC_complete, WAIS_Complete) %>%
  count()

```

## of complete WIAT, how many complete WISC?
```{r}
cog_df %>%
  filter(WIAT_Complete == 1) %>%
  group_by(WISC_complete) %>%
  count()

```

## questions: what do 2s mean for "complete" variables? 